---
title: "Help for FunctionDescriptiveStatisticsTable.R"
author: "Anna Schritz"
date: "14 janvier 2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG = "en")
```

## R function descr() and descrBy2()

R functions to get summary descriptive statistics for continuous and categorical variables together in one table/database.

### descr()
descr(var, name=var, database, print = TRUE)

Arguments:  
  * var (character): name of variable  
  * name (character): english name of variable  
  * database: name of dataset  
  * print (logical): print variable name per iteration (useful if you have an error and want to know where the error occurs)  
  
  
#### Examples:
```{r descr examples}
library(NHANES) # for NHANES dataset
library(purrr) # for map()

# save dataset
nhanes <- NHANES

# load descr() and descrBy2() functions
source("K:/CCMS Public/R_SAS good practices/R/01 - January 2020/FunctionDescriptiveStatisticsTable.R")

descr(var = "Age", name = "Age", database = nhanes, print = TRUE)
descr(var = "Race1", name = "Race", database = nhanes, print = FALSE)

# use function descr() with more than one variable
# define vectors with names of variables where descriptives statistics are needed
varNames <- c("Age", "Weight", "Race1")
OutpNames <- c("Age", "Weight", "Race")


# run for-loop over all arguments from varNames vector 
tab_baset <- c()
for(i in 1:length(varNames)){
  drow <- descr(var = varNames[i], name = OutpNames[i], database = nhanes, print = FALSE)
  tab_baset <- rbind(tab_baset, drow)
}
tab_baset
# --> easier ways to run function with multiple variables without for-loop possible

# as list object
sapply(varNames, FUN = descr, database = nhanes)

```

### descrBy2()
descrBy2(var, name=var, catVar, database, contTest = "wilcox", print = TRUE)  
  
Currently it is only possible to use a categorical variable (varVar) with 2 levels to categorise!  
  
  
Arguments:   
  * var (character): name of variable   
  * name (character): modified name of variable  
  * catVar (character): categorical variable to define subgroups  
  * database: name of dataset  
  * contTest (character): "wilcox" if two-sample Wilcoxon test (non-parametic) should be used or "ttest" if Student's t-test should be used (parametric)  
  * print (logical): print variable name per iteration  
  

#### Examples:
```{r descrBy2}
descrBy2(var = "Age", name = "Age", catVar = "Gender", contTest = "ttest", database = nhanes)
descrBy2(var = "Race1", name = "Race", catVar = "Gender", database = nhanes)

# merge of summary outputs from several variables to one table
# using t-test for continuous variables
tab_byGender <- c()
for(i in 1:length(varNames)){
  drow <- descrBy2(var = varNames[i], name = OutpNames[i], catVar = "Gender", contTest = "ttest", database = nhanes, print = FALSE)
  tab_byGender <- rbind(tab_byGender, drow)
}



# using tw-sample Wilcoxon test for continuous variables
tab_byGender_wil <- c()
for(i in 1:length(varNames)){
  drow <- descrBy2(var = varNames[i], name = OutpNames[i], catVar = "Gender", contTest = "wilcox", database = nhanes, print = FALSE)
  tab_byGender_wil <- rbind(tab_byGender_wil, drow)
}

```


